\documentclass[11pt]{article}
\usepackage{color}
\usepackage{amsmath,amsthm,amssymb,multirow,paralist}
\usepackage[margin=0.8in]{geometry}
\usepackage{hyperref}

\begin{document}

\begin{center}
{\Large \textbf{COM S 4740/5740: Introduction to Machine Learning}\\Homework \#3}\\

\linethickness{1mm}\line(1,0){498}

\begin{enumerate}
\item Please put required code files and report into a
compressed file ``HW\#\_FirstName\_LastName.zip''
\item Unlimited number of submissions are
allowed on Canvas and the latest one will be graded.
\item {\color{red} No later submission is accepted.}
\item Please read and follow submission instructions. No exception
will be made to accommodate incorrectly submitted files/reports.
\item All students are required to typeset their reports using
latex. Overleaf
(\url{https://www.overleaf.com/learn/latex/Tutorials}) can be a
good start.
\end{enumerate}

\linethickness{1mm}\line(1,0){498}

\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{enumerate}

\item (20 points) Consider the toy data set $\{([0, 0], -1),
([2, 2], -1), ([2, 0], +1)\}$. Set up the dual problem for
the toy data set. Then, solve the dual problem and compute
$\alpha^*$, the optimal Lagrange multipliers. (Note that there will
be three weights $\boldsymbol w = [w_0, w_1, w_2]$ by considering
the bias.)

\item (20 points) In a separable case, when a multiplier
$\alpha_i > 0$, its corresponding data point $(\boldsymbol x_i,
y_i)$ is on the boundary of the optimal separating hyperplane
with $y_i(\boldsymbol w^T \boldsymbol x_i) = 1$.

Show that the inverse is not True. Namely, it is possible that
$\alpha_i = 0$ and $(\boldsymbol x_i, y_i)$ is on the boundary
satisfying $y_i(\boldsymbol w^T \boldsymbol x_i) = 1$.

[Hint: Consider a toy data set with two positive examples at
([0,0], +1) and ([1, 0], +1), and one negative example at ([0,
1], -1).] (Note that there will be three weights $\boldsymbol w =
[w_0, w_1, w_2]$ by considering the bias.)

\item (20 points) \textbf{Kernel Function:} A function $K$
computes $K(\boldsymbol x_i, \boldsymbol x_j) =
-\boldsymbol{x_i}^T \boldsymbol x_j $. Is this function a valid
kernel function for SVM? Prove or disprove it.

\end{enumerate}

\end{document}
